---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# trickypdf

## Purpose

At a basic level, R users have efficient tools to extract text from pdf documents.
The [pdftools](https://CRAN.R-project.org/package=pdftools), and the [Rpoppler](https://CRAN.R-project.org/package=Rpoppler) packages are extremely useful and handy.

There are couple of issues that may arise when processing pdf that remain tricky. The *trickypdf*-package offers a class *PDF* to handle problems that reoccurringly cause headaches when processing pdf
documents. The class offers methods to...

  * remove stuff outside the main text region (page headers, page numbers etc) as a preprocessing step. 
  * handle multi-column layouts;
  * reconstruct lines of text, if the (OCRed) document has been scanned in tilted fashion;
  * reconstruct paragraphs.

The output will be a valid XML document, with optional document metadata. XML
is generated to serve as the input to a Natural Language Processing (NLP) pipeline. A method to 
create browsable html from the xmlified pdf document is meant to assist quality checking in 
corpus preparation. The package thus makes handling tricky problems with processing pdf
documents much easier!

## Installation

The package relies on the pdftohtml command line utility that is part of the poppler library.
On Linux, installation is easy:

```{sh, eval = FALSE}
sudo apt-get install -y poppler-utils
```

On MacOS, it is recommended to use Homebrew as a package manager. To install brew, open a terminal
and install Homebrew.

```{sh, eval = FALSE}
/usr/bin/ruby -e "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install)"
```

The snippet is taken from the [Homebrew website](https://brew.sh). You can then install poppler as 
follows,

```{sh, eval = FALSE}
brew install poppler
```

To check that poppler is installed, check the availability of the pdftohtml command line utility.

```{sh, eval = FALSE}
pdftohtml -v
```

The trickypdf-package imports from a set of standard R packages that are available from CRAN using 
`install.packages`.

```{r, eval = FALSE}
install.packages(
  pkgs = c("methods", "xml2", "pbapply", "Rpoppler", "htmltools", "markdown", "stringi", "plyr")
  )
```

Finally, trickypdf can be installed from GitHub using devtools.

```{r, eval = FALSE}
install.packages("devtools") # if you have not yet installed devtools
devtools::install_github("PolMine/trickypdf")
```


## Usage

```{r}
library(trickypdf)
```

### Scenario 1: PDF with column titles, page numbers, footnotes

As a sample document, the package includes the United Nations Millenium Declaration (pdf version).
We create an instance of the PDF class by supplying the filename.

```{r}
doc <- system.file(package = "trickypdf", "extdata", "pdf", "UN_Millenium_Declaration.pdf")
Decl <- PDF$new(filename_pdf = doc)
```

Let us have a look at the document. The *show_pdf*-method will open a browser window with the document.

```{r}
Decl$show_pdf()
```

For the pages of the document, we define boxes that define the area of the text that we want to
retain.

```{r}
Decl$add_box(page = 1, box = c(left = 110, top = 290, width = 400, height = 415))
Decl$add_box(page = 2, box = c(left = 110, top = 80, width = 400, height = 644))
Decl$add_box(page = 3, box = c(left = 110, top = 80, width = 400, height = 580))
Decl$add_box(page = 4, box = c(left = 110, top = 80, width = 400, height = 570))
Decl$add_box(page = 5, box = c(left = 110, top = 80, width = 400, height = 550))
Decl$add_box(page = 6, box = c(left = 110, top = 80, width = 400, height = 515))
Decl$add_box(page = 7, box = c(left = 110, top = 80, width = 400, height = 550))
Decl$add_box(page = 8, box = c(left = 110, top = 80, width = 400, height = 580))
Decl$add_box(page = 9, box = c(left = 110, top = 80, width = 400, height = 290))
```

We now drop anything outside the boxes (i.e page numbers, headers and column titles, footnotes). We
get the remaining text, do some postprocessing and xmlify things. By generating a html document, 
we perform a quality check.

```{r}
Decl$remove_unboxed_text_from_all_pages()
Decl$get_text_from_pages()
Decl$purge()
Decl$xmlify()
Decl$xml2html()
Decl$browse()
```

Looks ok? Let is save the file to disk.

```{r}
output <- tempfile(fileext = ".xml")
Decl$write(filename = output)
```


### Scenario 2: PDF with two columns. 

The second scenario may appear somewhat more difficult: It is a protocol of a meeting of
the UN General Assembly, a document with a two column layout.

```{r, eval = TRUE}
doc <- system.file(package = "trickypdf", "extdata", "pdf", "UN_GeneralAssembly_2016.pdf")
UN <- PDF$new(filename_pdf = doc)
UN$show_pdf()
```

As in the previous example, we define boxes. But this time, we define a second box for each
page that does not replace the previously defined box.

```{r}
UN$add_box(page = 1, box = c(top = 380, height = 250, left = 52, width = 255))
UN$add_box(page = 1, box = c(top = 232, height = 400, left = 303, width = 255), replace = FALSE)
UN$add_box(page = 2, box = c(top = 80, height = 595, left = 52, width = 255))
UN$add_box(page = 2, box = c(top = 80, height = 595, left = 303, width = 255), replace = FALSE)
```

In the first scenario, we dropped anything outside the boxes and got everything that was still on the
page. Now we explicitly grab the text in the boxes.

```{r}
UN$get_text_from_boxes(paragraphs = TRUE)
UN$xmlify()
```

Let us inspect the html of the resulting xml.

```{r}
UN$xml2html()
UN$browse()
```

Does it look ok? Let's save it to disk.

```{r}
output <- tempfile(fileext = ".xml")
UN$write(filename = output)
```


## Contributing to package development

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.